---
title: "Data Exploration for Museum Viz"
author: "Derya Akbaba, Madi Cooley, & Cole Polychronis"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

In this document, we explore the structure and contents of our cleaned data csv file, to understand how to effectively visualize it, as well as to identify any interesting datapoints to highlight.

## Preliminaries

Load necessary packages:

```{r, message = FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)
```

Load data for museum artifacts:

```{r}
museum_data_raw <- read.csv("cleaned-data.csv", header = TRUE)
museum_data <- data.frame(museum_data_raw)
museum_data$continent = substr(museum_data$continent, 1, 2)
str(museum_data)
```

## Exploration of Creation Dates

```{r}
museum_data_abs = subset(museum_data, created_date >= 0)
```

```{r}
ggplot(museum_data_abs, aes(x=created_date)) + geom_histogram(bins = 100) + labs(title = "Histogram of Creation Dates Faceted by Museum (After Common Era)") + facet_grid(museum ~ .) +
    theme(strip.text.y = element_text(angle = 0, hjust = 0))
```

From these histograms, we can see several interesting patterns or characteristics of the data. For one, we can see that both the Canada Science and Tech Museums and the MOMA have artifacts that are exclusively created within the last 200 years. We can also see that the collections for the Cooper-Hewitt and the Minneapolis Institute have relatively recently created collections, spanning back to about 500 years ago.

Additionally, we can see that with respect to the Cleveland Museum of Art, the Met, and the Penn Museum, while their collections are clustered closer to common day, they are in possession of artifacts that span back centuries and that it is likely these museums that posses artifacts that are dated to before the common era.

```{r}
museum_data_bc = subset(museum_data, created_date < 0)
```

```{r}
ggplot(museum_data_bc, aes(x=created_date)) + geom_histogram(bins = 100) + labs(title = "Histogram of Creation Dates Faceted by Museum (Before Common Era)") + facet_grid(museum ~ .) +
    theme(strip.text.y = element_text(angle = 0, hjust = 0))
```

Subsetting the data to include only artifacts that were developed before the Common Era (B.C. or B.C.E), we can see that it was in fact the three museums from earlier that possess artifacts dating back to before the Common Era. However, these seem to date back to roughly 10,000 B.C. or earlier, with only a few artifacts dating back to earlier. Below, we plot the histogram for artifacts that date back to 10,000 B.C. or earlier, with a table of the remaining artifacts that date back earlier.

```{r}
museum_data_bc_10000 = subset(museum_data, created_date < 0 & created_date > -10000)
```

```{r}
ggplot(museum_data_bc_10000, aes(x=created_date)) + geom_histogram(bins = 100) + labs(title = "Histogram of Creation Dates Faceted by Museum (Before Common Era)") + facet_grid(museum ~ .) +
    theme(strip.text.y = element_text(angle = 0, hjust = 0))
```

```{r}
museum_data_pre_10000 = subset(museum_data, created_date < -10000)
count(museum_data_pre_10000, "museum")
```

From this histogram and table, we can see that both the Met and the Penn museum artifacts that are B.C. were largely created about 2500 B.C. or earlier. Additionally, the Cleveland Museum has relatively few artifacts that are dated to before the Common Era. Finally, from the table, we can see that the Penn Museum has the vast majority of artifacts that date back to before 10,000 B.C., with the Met having a few as well.

## Exploration of Acquisition Dates

```{r}
ggplot(museum_data, aes(x=acquisition_date)) + geom_histogram(bins = 50) + labs(title = "Histogram of Acquisition Dates Faceted by Museum") + facet_grid(museum ~ .) +
    theme(strip.text.y = element_text(angle = 0, hjust = 0))
```

From these histograms of acquisition dates, we can see that there is a substantial amount of collection during the early 1900's, especially for the Met and the Penn Musuem. Additionally for these two musuems, we can see that there is specifically a large amount of collection in the decade leading up to the second World War. In this same vein, we can see that the Cleveland Museum and the Minneapolis Institute has a substantial amount of collection during the 1920's, which could possibly be due, in part, to the first World War. 

Moving on, we see that the Cleveland Museum, the Minneapolis Institute, and especially the MOMA acquired a large amount of artifacts during the 1960's and the early 1970's. The Met collected a lot of artifacts during the 1970's and early 1980's. Finally, and perhaps unsurpisingly, the Canada Science and Tech Museums and the MOMA collected a substantial amount of artifacts after the 2000's.

## Exploration of Continent

```{r}
ggplot(museum_data, aes(continent)) + geom_bar() + labs(title = "Number of artifacts from each Continent (aggregate)")
```

From this histogram, we can see that there is a clear bias towards collecting and presenting North American and European art. Additionally, we can see that there is a large amount of artifacts collected from Africa and Asia, which potentially could be a manifestation of the colonialist tradition. Before we explore this more, we first look to see if this trend holds for each museum. 

```{r}
count_museum_by_cont = data.frame(summarise(group_by(museum_data, museum, continent), count=n()))

ggplot(count_museum_by_cont, aes(fill=museum, y=count, x=continent)) + geom_bar(position="dodge", stat="identity") + labs(title = "Number of Artifacts from each Museum grouped by Continent")
```

From this grouped bar chart, we can see that by a significant margin, the Met has the greatest amount of artifacts from Africa and a substantial amount from Europe and North America. We can see that a close second for African artifacts is the Penn museum, which also boasts the greatest number of Asian artifacts. The MOMA has the vast majority of artifacts from both Europe and North America. Interestingly, the only substantial amount of artifacts from Oceania belong to the Met. We can explore how these look per museum in the stacked bar chart below.

```{r}
ggplot(count_museum_by_cont, aes(x = museum, y = count, fill = continent)) + 
    geom_bar(position = "fill",stat = "identity") +
    scale_y_continuous(labels = scales::percent_format()) +
    theme(axis.text.x = element_text(angle = 90)) + 
    scale_x_discrete(labels=c("canada", "cleveland", "cooper-hewitt", "metropolitan", "minneapolis", "moma", "penn")) +
    labs(title = "Percentage of Artifacts from each Continent seperated by Museum")
```

From this chart, we can see that for 4 out of 6 of the museums, roughly a third of the artifacts come from North America and for 3 out of 6 museums, roughly a third of artifacts come from Europe. Further, the Canadian collection of museums, the Cleveland Museum, the Cooper-Hewitt, and the Minneapolis museum are almost completely comprised of artifacts from either North America or Europe. Artifacts from Oceania or South Africa are not readily represented in any of the museums.

Interestingly, about a half of the artifacts in the Met and a quarter of the atrifacts in the MOMA are from Africa. Based on visual inspection, it seems that the Minneapolis Museum has the least diverse set of artifacts, as over 80% of their artifacts are from Europe and almost the entire rest of the set are from North America. The Met and the MOMA seem like they could potentially be the most diverse, though they both definitely skew towards more African and more Asian artifacts, respectively. 

## Brief Exploration of Country

To end our analysis, we briefly explore where the most common source of artifacts is. To this end, we subset our data to show the countries that have more than 5000 artifacts presented in museums. 

```{r}
count_museum_by_country = data.frame(summarise(group_by(museum_data, country_of_origin), count=n()))

subset(count_museum_by_country, count > 5000)
```

From this, we can see that over 70,000 artifacts come from the United States, over 44,000 come from Egypt, and over 23,000 come from France. Overall, we can see that, unsurprisingly, a large amount of artifacts are pulled from either prominent countries in the West, such as the United States and countries in Western Europe or countries considered to be "cradles of civiliazation", such as China or Egypt.